Jack Antonoff-ification of pop music

### 2023-04-04 outputs/paper/paper.qmd
cleaning it up a bit

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Michaela Drouillard
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: html
number-sections: true
bibliography: references.bib
---

@@

\["*No*?" I thought to myself\]

@@

Midnight was produced by Swift and her longtime collaborator Jack Antonoff, who is the subject of this essay, and, I suspect, the reason behind both my blacking out, and the stranger's "Exactly".

Antonoff is prolific, to the extent that it is starting to bother people. He has produced and co-produced music for Lorde, Taylor Swift, Florence and the Machine, The Chicks, Clairo, The 1975, Grimes, Zayn, Pink, Lana Del Rey, Olivia Rodrigo, Minions: The Rise of Gru Soundtrack, and more. All of this and his own bands too: Bleachers, and fun.

When I was getting brunch with a friend one Sunday and a Swift song came on, and she said "Jack Antonoff's gonna kill us all", we both laughed. What my friend was referring to, of course, is the fact that Jack Antonoff seems to be everywhere these days. As Andrew Marantz writes in the New Yorker: "When his band releases an album, the world responds politely. When he produces one by Lorde or Lana Del Rey or Taylor Swift, the world wobbles on its axis".

What I am trying to get to here is that I don't really mind Jack Antonoff's music. I kind of like it actually. But I still say his voice with the same tone that everyone else does. There's just something about him... this sense that he's everywhere... and that he's just such a nice guy... Why are we all so fixated on him? What is it about Antonoff music that is so Antonoff-y?

To find an answer to this question, I turned people, statistics, and the overlord of engineered virality itself -- the Spotify API.

@@

@@

2.  his relationship to music growing up (you can skim this, but there are some gems in there)

@@

4.  His tongue in cheek response to the media

    > .... \[it's just subjective taste\]. there's just nothing to engage with, so ive had to fabricate a reason for controversy. and what ive come up with is this. \*looks at phone\* the real reason iive been saying for years that i have a visceral reaction to jack antonoffs production style, the REAL reason is this \-- cause during the rollout cycle of the album midnights, i wanted to ascribe taylor swifts authorship to a man and diminish women's voices. so actually, i'm a sexist. so now thats out there, publsihed as facts, in the news. with all the other facts..."

@@

6.  Which was further doubled down on when I found a screenshot he took of his Tweet drafts in June 2021, one of them reading: "people often say to me:"caleb, what's wrong with jack antonoff?". don't ask questions"

@@

@@

This is pretty vague. Especially valence -- is the positivity related to lyricsI compared the valence scores of classical music tracks to see if this score is based on lyrics, or the audio itself. Jupiter's valence score is 326.78% higher than Mars. This not only indicates the valence is based on audio features other than lyrical content, but that it is capturing a difference between the two songs' profiles.

@@

and danceability as "how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity. A value of 0.0 is least danceable and 1.0 is most danceable". Neither of these definitions fully cover what metrics and thresholds go into something like "danceability". However, whatever these features capture is easier to compare across artists in visualizations than more granular details about bars, beats, duration, and tatums -- a word I just learned, which refers to "the lowest regular pulse train that a listener intuitively infers from the timing of perceived musical events". The full documentation for each variable included in this study is available in the Appendix.

@@

@eq-bayes seems useful, eh?

>>>>

---
title: "Jack Antonoff-ification of pop music"
subtitle: "My subtitle if needed"
author: 
  - Michaela Drouillard
thanks: "Code and data are available at: LINK."
date: today
abstract: "In this essay, I will be exploring Jack Antonoff's production style using data from the Spotify API. Drawing from an interview with Caleb Gamman, a guy who went viral for guessing which song on Taylor Swift's Midnights album was produced by Antonoff within seconds, I will delve into the specific characteristics that define an "Antonoff-y" song, and assess whether or not the audio features defined by Spotify can capture this. I plan to lasso Glen McDonald,the Spotify engineer behind these features, into explaining them to me. Finally, I will armchair a little bit about the cultural significance of Antonoff's music and the reasons behind his ubiquity in the music industry."
format: html
number-sections: true
bibliography: references.bib
---

@@

\["*No*?"\]

@@

Midnight was produced by Swift and her longtime collaborator Jack Antonoff, who is the subject of this paper, and, I suspect, the reason behind both my blacking out, and the stranger's "Exactly".

Antonoff is prolific, to the extent that it is starting to bother people. He has produced and co-produced music for Lorde, Taylor Swift, Florence and the Machine, The Chicks, Clairo, The 1975, Grimes, Zayn, Pink, Lana Del Rey, Olivia Rodrigo, the Minions: The Rise of Gru Soundtrack, and more. All of this and his own bands too (Bleachers, and fun.). When I was getting brunch with a friend one Sunday and a Swift song came on, and she said "Jack Antonoff's gonna kill us all", we both laughed. What my friend was referring to, of course, is the fact that Jack Antonoff seems to be everywhere these days. As Andrew Marantz writes in the New Yorker: "When his band releases an album, the world responds politely. When he produces one by Lorde or Lana Del Rey or Taylor Swift, the world wobbles on its axis".

What I am trying to get to here is that I don't actually really mind Jack Antonoff's music, I kind of like it, but I still say his voice with the same tone that everyone else does. There's just something about him... this sense that he's everywhere... and that he's just such a nice guy... Why do music people seem to be a little annoyed with him? What is it about Antonoff music that is so Antonoff-y? Is the Jack Antonoff-fication of pop something we should be worried about?

To find an answer to this question, I turned to statistics, people who know more than me about describing music, and the overlord of engineered virality itself -- the Spotify API.

@@

(you can skim the quote chunks)

@@

2.  his relationship to music growing up (some gems in here)

@@

4.  His tongue in cheek response to some of Twitter's negative response to his video

    > .... \[it's just subjective taste\]. there's just nothing to engage with, so ive had to fabricate a reason for controversy. and what ive come up with is this. \*looks at phone\* the real reason iive been saying for years that i have a visceral reaction to jack antonoffs production style, the REAL reason is this \-- cause during the rollout cycle of the album midnights, i wanted to ascribe taylor swifts authorship to a man and diminish women's voices. so actually, i'm a sexist. so now thats out there, publishhed as facts, in the news. with all the other facts..."

@@

6.  Which was refuted even harder when I found a screenshot he tweeted of his tweet drafts in June 2021, one of them reading: "people often say to me:"caleb, what's wrong with jack antonoff?". don't ask questions"

@@

\[\[He ended up saying a lot of interesting things, which i will transcribe and somehow make use of in this paper later\]

@@

This is pretty vague. Especially valence -- is the positivity related to lyrics, or something else? I compared the valence scores of classical music tracks to see if this score is based on lyrics, or the audio itself. Jupiter's valence score is 326.78% higher than Mars. This not only indicates the valence is based on audio features other than lyrical content, but that it is in fact capturing a difference between the two songs' profiles.

@@

@@
